---
title: "Twitter coverage of the Genome Informatics 2018 conference"
author: "Luke Zappia"
date: "`r Sys.time()`"
output: 
  html_document: 
    number_sections: yes
    toc: yes
    toc_float: yes
---

```{r knitr, include = FALSE}
DOCNAME <- tools::file_path_sans_ext(knitr::current_input())
knitr::opts_chunk$set(autodep        = TRUE,
                      cache          = FALSE,
                      cache.path     = paste0("cache/", DOCNAME, "/"),
                      cache.comments = TRUE,
                      echo           = FALSE,
                      error          = FALSE,
                      fig.align      = "center",
                      fig.path       = paste0("figures/", DOCNAME, "/"),
                      fig.width      = 10,
                      fig.height     = 8,
                      message        = FALSE,
                      warning        = FALSE)
```

```{r libraries, include = FALSE}
# Twitter
library("rtweet")

# Data manipulation
library("dplyr")
library("tidyr")
library("lubridate")
library("forcats")

# Text analysis
library("tidytext")

# Graphs
library("igraph")

# Visualisation
library("ggplot2")
library("ggrepel")
library("wordcloud")
library("ggraph")

# Presentation
library("knitr")

# Here
library("here")
```

```{r params}
hashtag <- "#gi2018"
conf_name <- "Genome Informatics 2018"
timezone <- "Europe/London"
days <- c("2018-09-17", "2018-09-18", "2018-09-19", "2018-09-20")
accent <- "skyblue3"
```

```{r load, include = FALSE}
theme_set(theme_light())

data_file <- here("data", paste0(DOCNAME, ".Rds"))

if (file.exists(data_file)) {
    tweets <- readRDS(data_file)
} else {
    tweets <- search_tweets(hashtag, 10000)
    saveRDS(tweets, data_file)
}
```

# Introduction
An analysis of tweets from the `r conf_name` conference.
`r nrow(tweets)` tweets were collected using the `rtweet` R package:

# Timeline

## Tweets by day

```{r tweets-by-day}
tweets %>% 
  mutate(date = as_date(created_at, tz = timezone)) %>% 
  count(date) %>% 
  ggplot(aes(date, n)) + geom_col(fill = accent) +  
    labs(x = "Date", y = "Tweets",
         title = paste(hashtag, "tweets per day")) + 
    theme(axis.text = element_text(size = 12),
          axis.title = element_text(size = 12))
```

## Tweets by day and time

Filtered for dates `r min(days)` - `r max(days)` in the `r timezone` timezone.

```{r tweets-by-day-hour}
tweets_days <- tweets %>% 
  mutate(datetime = as_datetime(created_at, tz = timezone),
         hour = hour(datetime)) %>% 
  group_by(date = as_date(datetime), hour) %>% 
  filter(date >= as_date(min(days)), date <= as_date(max(days)))

tweets_days %>%
    summarise(count = n()) %>% 
    ggplot(aes(hour, count)) +
    geom_col(fill = accent) +
    facet_grid(strftime(date, "%b %d") ~ .) + 
    labs(x = "Hour", y = "Tweets",
         title = paste(hashtag, "tweets by time of day")) + 
    theme(axis.text = element_text(size = 12),
          axis.title = element_text(size = 12))
```

# Users

## Top tweeters

```{r tweets-top-users}
tweets %>% 
  count(screen_name) %>% 
  arrange(desc(n)) %>%
  slice(1:20) %>% 
  ggplot(aes(reorder(screen_name, n), n)) +
    geom_col(fill = accent) +
    coord_flip() + 
    labs(x = "Screen Name", y = "Tweets",
         title = paste(hashtag, "tweets by user"),
         subtitle = "top 20 users") + 
    theme(axis.text = element_text(size = 12),
          axis.title = element_text(size = 12))
```

## Top original tweeters

```{r tweets-top-users-orig}
tweets %>%
    filter(is_retweet == FALSE) %>%
    count(screen_name) %>% 
    arrange(desc(n)) %>%
    slice(1:20) %>% 
    ggplot(aes(reorder(screen_name, n), n)) +
        geom_col(fill = accent) +
        coord_flip() + 
        labs(x = "Screen Name", y = "Tweets",
             title = paste(hashtag, "tweets by user"),
             subtitle = "top 20 users (no retweets)") + 
        theme(axis.text = element_text(size = 12),
              axis.title = element_text(size = 12))
```

## Retweet proportion

```{r tweets-retweet-prop}
plot_data <- tweets %>%
    group_by(screen_name) %>%
    summarise(orig = sum(!is_retweet),
              retweet = sum(is_retweet)) %>%
    mutate(total = orig + retweet)

ggplot(plot_data, aes(total, (orig / total) - (retweet / total))) +
        geom_hline(yintercept = 0, colour = "red") +
        geom_point(colour = accent) +
        geom_text_repel(data = top_n(plot_data, 10, total), 
                        aes(label = screen_name)) +
        labs(x = "Total tweets",
             y = "<<< more retweets : more original tweets >>>",
             title = paste(hashtag, "original tweets compared to retweets"),
             subtitle = "top 10 users labelled") + 
        theme(axis.text = element_text(size = 12),
              axis.title = element_text(size = 12))
```

## Top tweeters by time

```{r tweets-top-users-time}
top_users <- tweets %>%
    count(screen_name) %>% 
    arrange(desc(n)) %>%
    slice(1:5) %>%
    pull(screen_name) %>%
    fct_inorder()

tweets %>%
    filter(screen_name %in% top_users) %>%
    mutate(screen_name = factor(screen_name,
                                levels = levels(top_users))) %>%
    mutate(datetime = as_datetime(created_at, tz = timezone)) %>%
    ggplot(aes(datetime, 1, shape = is_retweet, colour = screen_name)) +
        geom_jitter(width = 0, height = 1) +
        scale_color_brewer(palette = "Set1", guide = FALSE) +
        facet_wrap(~ screen_name, ncol = 1) +
        labs(x = "Datetime",
             title = paste(hashtag, "top users timeline"),
             subtitle = "when the top 5 users tweeted") + 
        theme(axis.text = element_text(size = 12),
              axis.title = element_text(size = 12),
              axis.title.y = element_blank(),
              axis.text.y = element_blank(),
              axis.ticks.y = element_blank(),
              legend.position = "bottom")
```

## Top tweeters by day

### All tweets {.tabset}

```{r tweets-top-users-days, results = "hide"}
plot_list <- lapply(seq_along(days), function(ind) {
    tweets_days %>% 
        filter(date == as_date(days[ind])) %>%
        ungroup() %>%
        count(screen_name) %>% 
        arrange(desc(n)) %>%
        slice(1:20) %>% 
        ggplot(aes(reorder(screen_name, n), n)) +
            geom_col(fill = accent) +
            coord_flip() +
            labs(x = "Screen Name", y = "Tweets",
                 title = paste(hashtag, "tweets by user, Day", ind),
                 subtitle = "top 20 users") + 
            theme(axis.text = element_text(size = 12),
                  axis.title = element_text(size = 12))
})

src_list <- lapply(seq_along(plot_list), function(ind) {
    src <- c("#### Day {{ind}} {.unnumbered}",
             "```{r tweets-top-users-day{{ind}}}",
             "plot_list[[{{ind}}]]",
             "```",
             "")
    knit_expand(text = src)
})

out <- knit_child(text = unlist(src_list))
```

`r out`

### Original tweets {.tabset}

```{r tweets-top-users-orig-days, results = "hide"}
plot_list <- lapply(seq_along(days), function(ind) {
    tweets_days %>% 
        filter(date == as_date(days[ind])) %>%
        ungroup() %>%
        filter(is_retweet == FALSE) %>%
        count(screen_name) %>% 
        arrange(desc(n)) %>%
        slice(1:20) %>% 
        ggplot(aes(reorder(screen_name, n), n)) +
            geom_col(fill = accent) +
            coord_flip() +
            labs(x = "Screen Name", y = "Tweets",
                 title = paste(hashtag, "tweets by user, Day", ind),
                 subtitle = "top 20 users (no retweets)") + 
            theme(axis.text = element_text(size = 12),
                  axis.title = element_text(size = 12))
})

src_list <- lapply(seq_along(plot_list), function(ind) {
    src <- c("#### Day {{ind}} {.unnumbered}",
             "```{r tweets-top-users-orig-day{{ind}}}",
             "plot_list[[{{ind}}]]",
             "```",
             "")
    knit_expand(text = src)
})

out <- knit_child(text = unlist(src_list))
```

`r out`

## Sources

```{r tweets-top-sources}
tweets %>% 
  distinct(screen_name, source) %>%
  count(source) %>% 
  filter(n >= 5) %>% 
  ggplot(aes(reorder(source, n), n)) +
    geom_col(fill = accent) +
    coord_flip() + 
    labs(x = "Source", y = "Tweets",
         title = paste(hashtag, "tweets by source"),
         subtitle = "distinct(screen_name, source) with >= 5 tweets") + 
    theme(axis.text = element_text(size = 12),
          axis.title = element_text(size = 12))
```

# Networks

## Replies

The "replies network", composed from users who reply directly to one another, 
coloured by page rank.

```{r reply-network}
tweets_replies <- tweets %>% 
  filter(!is.na(reply_to_screen_name)) %>% 
  select(screen_name, reply_to_screen_name) %>% 
  graph.data.frame(directed = TRUE)

V(tweets_replies)$label <- V(tweets_replies)$name
V(tweets_replies)$id    <- V(tweets_replies)$name
V(tweets_replies)$pr    <- page_rank(tweets_replies)$vector

ggraph(tweets_replies, layout = "fr") +
    geom_edge_link(arrow = arrow(length = unit(4, 'mm')), 
                   end_cap = circle(1, 'mm'),
                   colour = "darkgrey") +
    geom_node_point(aes(colour = pr)) +
    geom_node_text(aes(label = label), colour = "skyblue", repel = TRUE) +
    viridis::scale_colour_viridis() +
    theme_graph() +
    theme(legend.position = "none")
```

## Mentions

The "mentions network", where users mention other users in their tweets.

```{r mentions-network}
tweets_mentions <- tweets %>% 
  filter(!is.na(mentions_screen_name)) %>% 
  select(screen_name, mentions_screen_name) %>% 
  unnest(mentions_screen_name) %>% 
  mutate(mentions_screen_name = strsplit(mentions_screen_name, " ")) %>% 
  unnest(mentions_screen_name) %>% 
  graph.data.frame()

V(tweets_mentions)$label <- V(tweets_mentions)$name
V(tweets_mentions)$id    <- V(tweets_mentions)$name
V(tweets_mentions)$pr    <- page_rank(tweets_mentions)$vector
V(tweets_mentions)$kcore <- coreness(tweets_mentions)

lo_kcore <- V(tweets_mentions)$kcore < 3

tweets_mentions <- delete_vertices(tweets_mentions, V(tweets_mentions)[lo_kcore])

ggraph(tweets_mentions, layout = "fr") +
    geom_edge_link(arrow = arrow(length = unit(2, 'mm')), 
                   end_cap = circle(1, 'mm'),
                   width = 0.1, colour = "darkgrey") +
    geom_node_point(aes(colour = pr)) +
    geom_node_text(aes(label = label, size = pr), colour = "blue") +
    viridis::scale_colour_viridis() +
    theme_graph() +
    theme(legend.position = "none")
```

# Retweets

## Retweet proportion

```{r is-retweet}
tweets %>% 
  count(is_retweet) %>% 
  ggplot(aes(is_retweet, n)) + geom_col(fill = accent) + 
    labs(x = "Is retweet", y = "Tweets",
         title = paste(hashtag, "tweets by retweet status")) + 
    theme(axis.text = element_text(size = 12),
          axis.title = element_text(size = 12))
```

## Retweet count

```{r retweet-count}
tweets %>% 
  ggplot(aes(retweet_count)) +
    geom_histogram(bins = max(tweets$retweet_count), fill = accent) +
    labs(x = "Retweet count", y = "Tweets",
         title = paste(hashtag, "distribution of retweets per tweet")) + 
    theme(axis.text = element_text(size = 12),
          axis.title = element_text(size = 12))
```

## Top retweets

```{r most-retweeted}
tweets %>% 
  filter(is.na(retweet_status_id)) %>% 
  select(screen_name, text, retweet_count) %>% 
  arrange(desc(retweet_count)) %>% 
  distinct() %>%
  slice(1:10) %>% 
  kable(format = "html")
```

# Favourites

## Favourite proportion

```{r has-favorite}
tweets %>% 
  mutate(has_favorite = ifelse(favorite_count > 0, TRUE, FALSE)) %>% 
  count(has_favorite) %>%
  ggplot(aes(has_favorite, n)) + geom_col(fill = accent) + 
    labs(x = "Has favorite", y = "Tweets",
         title = paste(hashtag, "tweets by favorited status")) + 
    theme(axis.text = element_text(size = 12),
          axis.title = element_text(size = 12))
```

## Favourite count

```{r favorite-count}
tweets %>% 
  ggplot(aes(favorite_count)) +
    geom_histogram(bins = max(tweets$favorite_count), fill = accent) +
    labs(x = "Favorite count", y = "Tweets",
         title = paste(hashtag, "distribution of favorites per tweet")) + 
    theme(axis.text = element_text(size = 12),
          axis.title = element_text(size = 12))
```

## Top favourites

```{r most-favorited}
tweets %>% 
  select(screen_name, text, favorite_count) %>% 
  arrange(desc(favorite_count)) %>% 
  distinct() %>%
  slice(1:10) %>% 
  kable(format = "html")
```

# Quotes

## Quote proportion

```{r is-quote}
tweets %>% 
  count(is_quote) %>% 
  ggplot(aes(is_quote, n)) +
    geom_col(fill = accent) + 
    labs(x = "Is quote", y = "Tweets",
         title = paste(hashtag, "tweets by quote status")) + 
    theme(axis.text = element_text(size = 12),
          axis.title = element_text(size = 12))
```

## Quote count

```{r quotes-count}
tweets %>% 
  filter(!is.na(quoted_status_id)) %>% 
  count(quoted_status_id) %>% 
  ggplot(aes(n)) + geom_histogram(bins = 10, fill = accent) +
    labs(x = "Quote count", y = "Tweets",
         title = paste(hashtag, "distribution of quotes per tweet")) + 
    scale_x_continuous(limits = c(0, 10), breaks = seq(0, 10, 2)) + 
    theme(axis.text = element_text(size = 12),
          axis.title = element_text(size = 12))
```

## Top quotes

```{r most-quoted}
tweets %>% 
  filter(!is.na(quoted_status_id)) %>% 
  count(quoted_status_id) %>% 
  filter(n > 1) %>% 
  arrange(desc(n)) %>% 
#  slice(1:10) %>% 
  inner_join(select(tweets, screen_name, quoted_status_id,
                    is_retweet, text)) %>% 
  filter(is_retweet == FALSE) %>% 
  select(screen_name, text, quote_count = n) %>%
  distinct() %>%
  slice(1:10) %>%
  kable(format = "html")
```

# Media

## Media count

```{r has-media}
tweets %>% 
  mutate(has_media = !is.na(media_url)) %>% 
  count(has_media) %>% 
  ggplot(aes(has_media, n)) +
    geom_col(fill = accent) + 
    labs(x = "Has media", y = "Tweets",
         title = paste(hashtag, "tweets by media status")) + 
    theme(axis.text = element_text(size = 12),
          axis.title = element_text(size = 12))
```

## Top media

```{r liked-media}
tweets_media <- tweets %>% 
  filter(!is.na(media_url)) %>% 
  arrange(desc(favorite_count)) %>%
  filter(favorite_count > 0)

tweets_media %>%
  slice(1:10) %>% 
  select(screen_name, text, favorite_count) %>%
  kable(format = "html")
```

### Most liked media image

![](`r tweets_media[1, "media_url"]`)

# Tweet text

The top 100 words used 3 or more times.

```{r count-words}
data("stop_words")

tweets %>% 
  filter(is_retweet == FALSE) %>% 
  unnest_tokens(word, text) %>% 
  select(word) %>% 
  filter(!word %in% c(gsub("#", "", hashtag), "https", "t.co", "amp"),
         !word %in% tolower(tweets$screen_name), 
         !grepl("^\\d+$", word)) %>% 
  anti_join(stop_words) %>% 
  count(word) %>% 
  with(wordcloud(word, n, max.words = 100, min.freq = 3,
                 colors = brewer.pal(6, "Spectral")))
```
